---
title: "CL_Analysis_All"
author: "Jonas BÉAL"
date: "09/02/2020"
output:
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
#Infos from simulations
model <- c("BRAF_Model")
models_outputs <- c("Proliferation_b1","Proliferation_b2")

sim_cases <- c("mut","RNA", "mut_RNA")

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#Packages
#devtools::install_github("teunbrand/ggh4x")
if (!require("pacman")) install.packages("pacman")
list.of.packages <- c("readxl","knitr","corrplot","ggfortify","reshape2","gridExtra","magrittr", "tictoc", "ggrepel", "broom", "nlme", "lme4", "lmerTest", "eulerr", "plotly", "paletteer", "ggpubr", "cowplot", "ggcorrplot", "ggh4x", "tidyverse", "UpSetR", "latex2exp")
pacman::p_load(list.of.packages, character.only = TRUE)
select <- dplyr::select
tbl_to_df <- function(x){
  x <- x %>% as.data.frame %>%  remove_rownames %>% column_to_rownames(var="model_name")
}

#Paths
original_path <- normalizePath(getwd())
base_folder <- original_path %>% dirname %>% dirname
knitr::opts_knit$set(root.dir = base_folder)

#Colors
#paletteer_d("dutchmasters::anatomy")
colors_types <- c(CM="#B47562FF", CRC="#332826FF")
colors_perso <- c(Mutants="#2E3440FF", Rates="#4C566AFF")
```

# Import simulations results and screening data

We study the clinical value of some logical models with cell lines data from CellModelPassports portal and GDSC dataset. Let's first import simulation results and drug/CRISPR screening files.

```{r DataImport, echo=FALSE, message=FALSE, warning=FALSE}
#Genes involved in PC model
genenames <- read.table(paste0("Models/",model,"/",model,"_namesToHugo_curated.txt"),
                         header=T,sep="\t")
geneindex <- strsplit(as.character(genenames[,2]), split = ",") %>% sapply(function(l){gsub(" ","",l)})
geneindex <- data.frame(V1 = rep(genenames[,1], sapply(geneindex, length)), V2 = unlist(geneindex))
model_genes_HUGO <- unique(geneindex$V2) %>% sub("^\\s+", "", .)


#First import CL clinical data for further use to correlate with logical modelling results
CL_clin <- read_csv("Data/CL/Infos/model_list_20200115.csv") %>%
  select(model_id, model_name, synonyms, model_type, COSMIC_ID, BROAD_ID,
         contains("_data"), contains("tissue"), contains("cancer")) %>%
  filter(mutation_data | cnv_data | expression_data) %>%
  mutate(TCGA_label=if_else(cancer_type=="Colorectal Carcinoma", "CRC",
                            if_else(cancer_type=="Melanoma", "CM", cancer_type)))

#Focus on two cancer types only
good_labels <- c("CRC", "CM")
CL_clin %<>% filter(TCGA_label %in% good_labels)
my_cell_lines <- unlist(CL_clin$model_name)
my_cell_lines_s <- unlist(CL_clin$model_id)

#Fonction to transpose some tibbles
tibble_transpose <- function(df_input){
  df_output <- df_input %>% gather(var, value, -gene_symbol) %>%
    spread(gene_symbol, value) %>%
    rename(model_name=var) %>% 
    type_convert
}

#Import CL mutations profile
CL_mut_profile <- read_csv(paste0("Results/Profiles/",model,"_CL_mut.csv")) %>% rename(Patient_ID=X1) %>%
  select(-matches("_b2")) %>%
  rename_all(function(x) gsub("_b1", "", x)) %>%
  mutate(Label_Mut=NA_character_)
gene_names <- colnames(CL_mut_profile)
for (r in 1:nrow(CL_mut_profile)){
  interm <- "Mutations->"
  for (c in 2:(ncol(CL_mut_profile)-1)){
    if (!is.na(CL_mut_profile[r,c])){
      interm <- paste0(interm, gene_names[c],":",CL_mut_profile[r,c],",")
    }
  }
  if (interm!="Mutations->"){
    interm <- substr(interm, 1, nchar(interm)-1)
    } else {interm="Mutations:None"}
  CL_mut_profile[r,ncol(CL_mut_profile)] <- interm
}


#Import CL RNA profile
CL_RNA_profile <- read_csv(paste0("Results/Profiles/",model,"_CL_RNA_norm.csv")) %>% rename(Patient_ID=X1) %>%
  select(-matches("_b2")) %>%
  rename_all(function(x) gsub("_b1", "", x)) %>%
  select(Patient_ID,
         #SOX10, HGF, EGF, FGF, 
         ERK,
         #SPRY,
         p70, PDK1, CRAF) %>%
  mutate_at(vars(-Patient_ID), function(x) round(x, digits=2)) %>%
  mutate(Label_RNA=NA_character_)
gene_names <- colnames(CL_RNA_profile)
for (r in 1:nrow(CL_RNA_profile)){
  interm <- "RNA->"
  for (c in 2:(ncol(CL_RNA_profile)-1)){
    interm <- paste0(interm, gene_names[c],":",CL_RNA_profile[r,c],",")
  }
  CL_RNA_profile[r,ncol(CL_RNA_profile)] <- interm
}

#Import GDSC drug screening
Drugs_Profile_Conversion <- read_csv(paste0("Results/Profiles/", model, "_drug_conversion.csv")) %>%
  rename(Drug_ID=DRUG_ID, Drug_Name=DRUG_NAME) %>%
  mutate(Drug_ID=as.character(Drug_ID))

#Prepare conversion
DC <- Drugs_Profile_Conversion %>%
  mutate(Variant=1)
for (r in 2:nrow(DC)){
  prof <- DC$Drug_Target[r]
  DC[r,"Variant"] <- 1+sum(DC$Drug_Target[1:(r-1)]==prof)
}
DC %<>% mutate(Variant=paste0(Drug_Target, "_",Variant)) %>%
  add_row(Drug_ID=NA,Drug_Target="Zero",
          Drug_Name="Zero", Variant="Zero") %>%
  rename(Target=Drug_Target)


#BRAF_drugs <- c("PLX-4720", "Dabrafenib")
CL_Drugs <- read_xlsx("Data/CL/GDSC2_fitted_dose_response_15Oct19.xlsx") %>%
  filter(SANGER_MODEL_ID %in% my_cell_lines_s) %>%
  mutate(CELL_LINE_NAME=if_else(CELL_LINE_NAME=="LS-1034", "LS1034", CELL_LINE_NAME)) %>%
  rename(Patient_ID=CELL_LINE_NAME, Drug_Name=DRUG_NAME) %>%
  left_join(rename(CL_clin, Patient_ID=model_name) %>% select(Patient_ID, TCGA_label),
            by="Patient_ID") %>%
  filter(Drug_Name %in% DC$Drug_Name) %>%
  select(Patient_ID, Drug_Name, LN_IC50, AUC, TCGA_label) %>%
  left_join(select(DC, -Drug_ID), by="Drug_Name") %>%
  pivot_longer(one_of("LN_IC50", "AUC"), names_to="Metric", values_to="Value")
  
#Import CRISPR/Cas9 screening
CL_CC1 <- read_tsv("Data/CL/03_scaledBayesianFactors.tsv") %>%
  filter(Gene %in% model_genes_HUGO) %>%
  rename(LS1034=`LS-1034`) %>%
  select(Gene, one_of(my_cell_lines)) %>%
  pivot_longer(-Gene, names_to="Patient_ID", values_to="Value") %>%
  left_join(rename(CL_clin, Patient_ID=model_name) %>% select(Patient_ID, TCGA_label),
            by="Patient_ID")

CL_CC2 <- read_tsv("Data/CL/scaledBayesianFactors.tsv") %>%
  filter(Gene %in% model_genes_HUGO) %>%
  pivot_longer(-Gene, names_to="Patient_ID", values_to="Value") %>%
  rename(BROAD_ID=Patient_ID) %>%
  left_join(select(CL_clin, BROAD_ID, model_name, TCGA_label),
            by="BROAD_ID")  %>%
  mutate(model_name=if_else(model_name=="LS-1034", "LS1034", model_name)) %>%
  rename(Patient_ID=model_name) %>%
  filter(Patient_ID %in% my_cell_lines) %>%
  select(-BROAD_ID)

CL_CC_All <- bind_rows(mutate(CL_CC1, Metric="CC1"),
                       mutate(CL_CC2, Metric="CC2")) %>%
  rename(Variant=Gene) %>%
  left_join(rename(geneindex, Target=V1, Variant=V2),
            by="Variant")

#Merge all metric
CL_Metrics <- bind_rows(mutate(CL_Drugs, Screening="Drug"),
                        mutate(CL_CC_All, Screening="CRISPR"))
  
CL_Metrics_Sample_Size <- CL_Metrics %>%
  filter(!is.na(Value)) %>%
  group_by(TCGA_label, Screening, Metric, Variant) %>%
  summarize(Size=n()) %>%
  ungroup()

#Nodes data
import_ALL <- function(sim_case){
  simulations <- read_table2(paste0("Results/Simulations/",model,"_CL_",sim_case,".txt")) %>%
    select(-Time, -n_profile) %>%
    mutate(Proliferation=(Proliferation_b1 + Proliferation_b2)/2)
  return(simulations)
}

simulations <- sapply(sim_cases, import_ALL, simplify = F, USE.NAMES = T)


print("All imports OK")
```

# Clinical characterization of cell lines

## Different screenings and metrics

But first, is there any consistence between all these values?

```{r plot_correlation_metrics, eval=TRUE, echo=FALSE, fig.height=7, fig.width=9, message=FALSE, warning=FALSE}
CL_Metrics_Reduced <- filter(CL_Metrics, Metric %in% c("AUC", "CC2")) %>%
  mutate(Label=paste0(Screening, "_", Variant))

corr <- select(CL_Metrics_Reduced, Patient_ID, Label, Value) %>%
  rename(Label_y=Label, Value_y=Value) %>%
  left_join(select(CL_Metrics_Reduced, Patient_ID, Label, Value),
            by="Patient_ID") %>%
  group_by(Label, Label_y) %>%
  summarise(Spearman=cor(Value, Value_y,
                         method="pearson", use="pairwise.complete.obs"),
            Spearman_t=cor.test(Value, Value_y,
                         method="pearson", use="pairwise.complete.obs")$p.value,
            N=n()) %>%
  mutate(Screening=if_else(grepl("CRISPR_", Label), "CRISPR", "Drug") %>%
           factor(levels=c("CRISPR", "Drug")),
         Screening_y=if_else(grepl("CRISPR_", Label_y), "CRISPR", "Drug")%>%
           factor(levels=c("Drug", "CRISPR")))

ggplot(corr, aes(x=Label, y=Label_y)) +
  geom_tile(aes(fill=Spearman)) +
  scale_fill_paletteer_c("pals::ocean.balance", limits=c(-1, 1)) +
  theme_pubclean() +
  theme(axis.title.x = element_blank(),
        legend.position = "right",
        axis.text.x = element_text(angle = 90)) +
  labs(fill="Correlation",
       y="Treatments",
       title="Pearson correlations between outcomes")

```

# Data reprocessing

We define a normalised variable based on level without any drug inhibition (*i.e* $Proliferation_{normalised} = Proliferation_{withDrug} / Proliferation_{withoutDrug}$)


```{r reprocess, echo=FALSE, message=FALSE, warning=FALSE}

#Merge all simulation cases
data_all <- simulations %>% melt %>% 
  filter(variable=="Proliferation") %>%
  spread(key="variable", value="value") %>%
  rename(SimCase=L1)

#Normalize
data_all_z <- filter(data_all, Drug_ID=="Zero") %>%
  select(-Drug_ID) %>%
  rename_at(vars(-Patient_ID, -SimCase), function(x) paste0(x, "_noDrug"))
data_all %<>%
  rename_at(vars(-Patient_ID, -Drug_ID, -SimCase), function(x) paste0(x, "_Drug")) %>%
  left_join(data_all_z, by = c("Patient_ID", "SimCase")) %>%
  #mutate(Proliferation_Norm=Proliferation_Drug-Proliferation_noDrug) %>%
  mutate(Proliferation_Norm=Proliferation_Drug/Proliferation_noDrug) %>%
  filter(Drug_ID!="Zero")

#Merge with clinical data
data_all_clin <- data_all %>%
  rename(Target=Drug_ID) %>%
  left_join(CL_Metrics,
            by=c("Patient_ID", "Target")) %>%
  left_join(select(CL_mut_profile, Patient_ID, Label_Mut)) %>%
  left_join(select(CL_RNA_profile, Patient_ID, Label_RNA)) %>%
  pivot_longer(cols=starts_with("Proliferation_"),
               names_to="Output", values_to="Score") %>%
  filter(!is.na(Value)) %>%
  mutate(P_Mutants=case_when(
           SimCase %in% c("mut", "mut_RNA") ~ "Mutations",
           SimCase %in% c("RNA") ~ " "
         ),
         P_Rates=case_when(
           SimCase %in% c("mut") ~ " ",
           SimCase %in% c("RNA",  "mut_RNA") ~ "RNA"
         ),
         SimCase=factor(SimCase, levels=sim_cases)) %>%
  mutate(Output=case_when(
    Output=="Proliferation_Drug" ~ "Prolif. w/ drug",
    Output=="Proliferation_noDrug" ~ "Prolif. w/o drug",
    Output=="Proliferation_Norm" ~ "Prolif. norm."
  ) %>%
    factor(levels=c("Prolif. w/o drug", "Prolif. w/ drug", "Prolif. norm.")))


```


# Simulations: a first quantitative approach with correlations

```{r correlations, echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}

corr <- group_by(data_all_clin, SimCase, P_Mutants, P_Rates,
                 Target, Screening, Metric, Variant, Output) %>%
  summarise(Number=n(),
            Spearman=cor(Value, Score,
                         method="spearman", use="pairwise.complete.obs"),
            Spearman_t=cor.test(Value, Score,
                         method="spearman", use="pairwise.complete.obs")$p.value,
            Pearson=cor(Value, Score,
                        method="pearson", use="pairwise.complete.obs"),
            Pearson_t=cor.test(Value, Score,
                        method="pearson", use="pairwise.complete.obs")$p.value) %>%
  ungroup %>%
  mutate(Label=if_else(Pearson_t<0.05, 
                       paste0(round(Pearson, digits=2), "\n(",
                              case_when(
                                Pearson_t<0.001~"***",
                                Pearson_t<0.01~"**",
                                Pearson_t<0.05~"*"), ")"),
                       ""))


#Now the plot function
overall_corr <- function(corr, target,
                         output=c("Prolif. w/o drug","Prolif. w/ drug", "Prolif. norm."),
                         sc_red=sim_cases){
  
  corr_red <- filter(corr,
                     SimCase %in% sc_red,
                     Target==UQ(target),
                     Output %in% UQ(output))
  
  p_table <- select(corr_red, SimCase, P_Mutants, P_Rates) %>%
    pivot_longer(starts_with("P_"), names_to="P_Method", values_to="P_Type") %>%
    mutate(Alpha_Status=if_else(P_Type==" ", 0, 1),
           P_Method=if_else(P_Method=="P_Mutants", "Discrete", "Continuous") %>%
             factor(levels=c("Discrete", "Continuous"))) %>%
    distinct %>%
    ggplot(aes(x=P_Method, y=SimCase)) +
    geom_tile(aes(alpha=as.factor(Alpha_Status)#,
                  ),
              color="black", size=0.5,fill="grey85",
              show.legend = FALSE) +
    geom_text(aes(label=P_Type), color="black") +
    theme_pubclean() +
    theme(axis.ticks.y = element_blank(),
          axis.text.y = element_blank()) +
    scale_fill_manual(values=colors_perso) +
    scale_alpha_manual(values = c(0.4,1)) +
    labs(x="Personalization methods",
         y="Combinations of personalization methods"
         )

p_drugs <- filter(corr_red, Screening=="Drug", Metric=="AUC") %>%
  ggplot(aes(x=1, y=SimCase)) +
  geom_tile(aes(fill=Pearson)) +
  geom_text(aes(label=Label), size=3) +
  facet_nested(.~Output+Variant) +
  scale_fill_paletteer_c("pals::ocean.balance", limits=c(-1, 1))+
  theme_pubclean() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "right") +
  guides(fill=FALSE)

p_crispr <- filter(corr_red, Screening=="CRISPR") %>%
  ggplot(aes(x=Metric, y=SimCase)) +
  geom_tile(aes(fill=Pearson)) +
  geom_text(aes(label=Label), size=3) +
  facet_nested(.~Output+Variant) +
  scale_fill_paletteer_c("pals::ocean.balance", limits=c(-1, 1))+
  theme_pubclean() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "right") +
  labs(fill="Correlation")

if (nrow(filter(corr_red, Screening=="Drug"))>0){
  egg::ggarrange(p_table, p_drugs, p_crispr, nrow = 1, widths = c(2,4,4),
               labels = c("A", "B", "C"))
  
  egg::ggarrange(p_table, p_drugs, p_crispr, nrow = 1, widths = c(2,4,4),
               labels = c("A", "B", "C"))
} else {
  egg::ggarrange(p_table, p_crispr, nrow = 1, widths = c(2,4),
               labels = c("A", "C"))
  }
}

overall_corr(corr, "BRAF",
             output=c("Prolif. w/o drug","Prolif. w/ drug", "Prolif. norm."))

```

Some interesting points:

  * We observe very significant correlations for both drug and CRISPR screening
  * Normalised proliferation appeared as the best proxy for drug sensitivity
  * RNA alone is usually not significant but its conjunction with mutation usually results in better results
  
Here are additional plots for other targets;
  
```{r correlations2, echo=FALSE, fig.height=6, fig.width=9, message=FALSE, warning=FALSE}
overall_corr(corr, "p53",
             output=c("Prolif. norm."))

overall_corr(corr, "PI3K",
             output=c("Prolif. norm."))
```
  
For drugs we will focus on PLX (but very similar to PLX in all aspects, no particular criterion to distinguish) and AUC metric (less sensitive to extrapolation). For CRISPR screening we will focus on CC2 dataset, more balanced in CM and CRC. For output we will also focus on normalisedx Proliferation scores:

Here is the pruned version of the plot for publication:

```{r correlations_pub, echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE, eval=TRUE}

corr_red <- filter(corr,
                   SimCase %in% sim_cases,
                   Target=="BRAF",
                   Output=="Prolif. norm.") %>%
  mutate(SimCase=factor(SimCase, levels=rev(sim_cases)))
  
p_table <- select(corr_red, SimCase, P_Mutants, P_Rates) %>%
    pivot_longer(starts_with("P_"), names_to="P_Method", values_to="P_Type") %>%
    mutate(Alpha_Status=if_else(P_Type==" ", 0, 1),
           P_Method=if_else(P_Method=="P_Mutants", "Strict NV", "Soft NV") %>%
             factor(levels=c("Strict NV", "Soft NV")),
           Title="Personalization methods") %>%
    distinct %>%
    ggplot(aes(x=P_Method, y=SimCase)) +
    geom_tile(aes(alpha=as.factor(Alpha_Status)#,
                  #fill=P_Method
                  ),
              color="black", size=0.5,fill="grey85",
              show.legend = FALSE) +
    geom_text(aes(label=P_Type), color="black") +
    theme_pubclean() +
    theme(axis.ticks.y = element_blank(),
          axis.text.y = element_blank()) +
    scale_fill_manual(values=colors_perso) +
    scale_alpha_manual(values = c(0.4,1)) +
    labs(x="Personalization methods",
         y="Combinations of personalization methods"#,
         #title="Omics type and personalization"
         )

p_table2 <- select(corr_red, SimCase, P_Mutants, P_Rates) %>%
    pivot_longer(starts_with("P_"), names_to="P_Method", values_to="P_Type") %>%
    mutate(Alpha_Status=if_else(P_Type==" ", 0, 1),
           P_Method=if_else(P_Method=="P_Mutants", "Discrete", "Continuous") %>%
               factor(levels=c("Discrete", "Continuous")),
           Title="Personalization methods") %>%
    distinct %>%
    ggplot(aes(x=1, y=SimCase)) +
    geom_tile(aes(alpha=as.factor(Alpha_Status)#,
                  #fill=P_Method
    ),
    color="black", size=0.5,fill="grey85",
    show.legend = FALSE) +
    geom_text(aes(label=P_Type), color="black") +
    facet_nested(.~Title+P_Method) +
    theme_pubclean() +
    theme(axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_text(size = 12)) +
    scale_fill_manual(values=colors_perso) +
    scale_alpha_manual(values = c(0.2,1)) +
    labs(y="Data types and\npersonalization methods"#,
         #title="Omics type and personalization"
    )

p_drugs <- filter(corr_red, Screening=="Drug", Metric=="AUC") %>%
  mutate(Output="BRAF inhibition: Drugs",
         Variant=if_else(Variant=="BRAF_1", "PLX-4720", "Dabrafenib") %>%
           factor(levels=c("PLX-4720", "Dabrafenib"))) %>%
  ggplot(aes(x=1, y=SimCase)) +
  geom_tile(aes(fill=Pearson)) +
  geom_text(aes(label=Label), size=5,fontface = "bold") +
  facet_nested(.~Output+Variant) +
  scale_fill_paletteer_c("pals::ocean.balance", limits=c(-1, 1))+
  theme_pubclean() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 12),
        legend.position = "right") +
  guides(fill=FALSE)

p_crispr <- filter(corr_red, Screening=="CRISPR") %>%
  mutate(Output="BRAF inhibition: CRISPR/Cas9",
         Metric=if_else(Metric=="CC1", "Sanger data", "Broad data") %>%
           factor(levels=c("Broad data", "Sanger data"))) %>%
  ggplot(aes(x=1, y=SimCase)) +
  geom_tile(aes(fill=Pearson)) +
  geom_text(aes(label=Label), size=5, fontface = "bold") +
  facet_nested(.~Output+Metric) +
  scale_fill_paletteer_c("pals::ocean.balance", limits=c(-1, 1))+
  theme_pubclean() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 14),
        legend.position = "right",
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
  labs(fill="Correlation")

p_1 <- egg::ggarrange(p_table2, p_drugs, p_crispr, nrow = 1, widths = c(2,4,4)#,
               #top = "Correlations with drug screening",
               #labels = c("A", "B", "C")
               )

#p_1
```

# Explore the results with scatter plots

## Simple plots

Now we want to have a deeper understanding of these correlation relations looking at the scatter plots

```{r scatter, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.height=6, fig.width=9}

data_all_clin_reduced <- filter(data_all_clin,
                                Metric %in% c("CC2", "AUC"),
                                !is.na(Value)) %>%
  mutate(Label=paste0(Patient_ID, " (", TCGA_label,")\n",Label_Mut, "\n", Label_RNA))

#Plot function
scatter_plot <- function(data_all_clin_reduced, target,
                         output=c("Prolif. norm."),
                         sc_red=sim_cases){
  
  plot_data <- filter(data_all_clin_reduced,
                      Target==UQ(target),
                      Output %in% output,
                      SimCase %in% sc_red) %>%
    left_join(select(CL_mut_profile, Patient_ID, BRAF),
              by = "Patient_ID") %>%
    mutate(BRAF=case_when(
      BRAF==1 ~ "Gain of function",
      BRAF==0 ~ "Loss of function",
      is.na(BRAF) ~ "No mutation"
    ))
  
  p_table_red <- select(plot_data, SimCase, P_Mutants, P_Rates) %>%
   pivot_longer(starts_with("P_"), names_to="P_Method", values_to="P_Type") %>%
   distinct %>%
   mutate(Alpha_Status=if_else(P_Type==" ", 0, 1),
         P_Method=if_else(P_Method=="P_Mutants", "Mutants", "Rates")) %>%
   ggplot(aes(x=P_Method, y=1)) +
      geom_tile(aes(alpha=as.factor(Alpha_Status)),
                color="black", size=0.5,fill="grey85",
                show.legend = FALSE) +
      geom_text(aes(label=P_Type), color="black", angle=90) +
      theme_pubclean() +
      theme(axis.ticks.y = element_blank(),
            axis.text.y = element_blank(),
            strip.background = element_blank(),
            strip.text = element_blank(),
            panel.grid.major.y = element_blank()) +
      scale_fill_manual(values=colors_perso) +
      scale_alpha_manual(values = c(0.4,1)) +
      labs(x="Personalization methods",
           y="Combinations of personalization methods") +
      facet_grid(SimCase~.)
  
  p_drugs <- filter(plot_data, Screening=="Drug") %>%
    ggplot(aes(x=Score, y=Value)) +
    geom_point(aes(color=TCGA_label), show.legend = FALSE) +
    facet_nested(SimCase~Output+Drug_Name) +
    theme_pubclean() +
    scale_color_manual(values = colors_types)  +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          strip.background.y = element_blank(),
          strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7)) +
    labs(y="Drug screening metric (AUC)")
  
  p_crispr <- filter(plot_data, Screening=="CRISPR") %>%
    ggplot(aes(x=Score, y=Value)) +
    geom_point(aes(color=TCGA_label), show.legend = FALSE) +
    facet_nested(SimCase~Output+Variant) +
    theme_pubclean() +
    scale_color_manual(values = colors_types)  +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          strip.background.y = element_blank(),
          strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7)) +
    labs(y="CRISPR metric (Bayes)")
  
  egg::ggarrange(p_table_red, p_drugs, p_crispr,
                 nrow = 1, widths = c(2,4,4),
                 top = target)

}

scatter_plot(data_all_clin_reduced, "BRAF", output=c("Prolif. norm."))
```

Here is the version for publication:

```{r scatter_pub, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.height=6, fig.width=9}
colors_types_4 <- c("CM (BRAF gain of function)"="#B47562FF",
                  "CM (no BRAF mutations)"="#DCB6AAFF",
                  "CRC (BRAF gain of function)"="#332826FF",
                  "CRC (no BRAF mutations)"="#9A9494FF")

plot_data <- filter(data_all_clin,
                    Target=="BRAF",
                    Metric %in% c("CC2", "AUC"),
                    Drug_Name!="Dabrafenib" | Screening=="CRISPR",
                    Output=="Prolif. norm.",
                    SimCase %in% sim_cases,
                    !is.na(Value)) %>%
  left_join(select(CL_mut_profile, Patient_ID, BRAF),
              by = "Patient_ID") %>%
  mutate(Label=case_when(
    BRAF==1 & TCGA_label=="CM" ~ "CM (BRAF gain of function)",
    is.na(BRAF) & TCGA_label=="CM" ~ "CM (no BRAF mutations)",
    BRAF==1 & TCGA_label=="CRC" ~ "CRC (BRAF gain of function)",
    is.na(BRAF) & TCGA_label=="CRC" ~ "CRC (no BRAF mutations)"
  )) %>%
    mutate(BRAF=case_when(
      BRAF==1 ~ "Gain function",
      BRAF==0 ~ "Loss function",
      is.na(BRAF) ~ "No mutation"
    ),
    SimCase=case_when(
      SimCase=="mut" ~ "Mutations",
      SimCase=="RNA" ~ "RNA",
      SimCase=="mut_RNA" ~ "Mutations & RNA"
    ) %>% factor(levels=c("Mutations", "RNA", "Mutations & RNA"))
    )
  
  p_drugs_scatter2 <- filter(plot_data, Screening=="Drug") %>%
    mutate(Label_Void=" ") %>%
    ggplot(aes(x=Score, y=Value)) +
    geom_label_repel(aes(label=Label_Void, fill=Label),
                     show.legend = FALSE,
                     box.padding = 0, label.padding = 0.02,
                     segment.color = NA) +
    stat_cor(aes(label = ..r.label..), method="pearson", label.y.npc="top") +
    facet_grid(SimCase~Drug_Name) +
    theme_pubclean() +
    scale_fill_manual(values = colors_types_4) +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 14),
          strip.background.y = element_blank(),
          strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7),
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
    labs(#x=expression(paste("Proliferation"[BRAFinhibition]," / Proliferation"[control]))#,
         #x=TeX("$\\frac{Proliferation_{BRAFinhibited}}{Proliferation_{control}}$"),
         y="Drug screening metric\n(AUC)"
         )
  
  plot_data2 <- filter(plot_data, Screening=="CRISPR") %>%
    mutate(Variant="Broad data", Label_Void=" ",
           Label_Select = if_else(Patient_ID %in% c("SK-MEL-24", "COLO-678", "HT-29", "HT-144", "HT55", "SK-MEL-30", "SK-MEL-2", "UACC-62"),
                                  Patient_ID, " "))
    p_crispr_scatter2 <- ggplot(plot_data2, aes(x=Score, y=Value)) +
    geom_label_repel(aes(label=Label_Void, fill=Label),
                     show.legend = TRUE,
                     box.padding = 0, label.padding = 0.02,
                     segment.color = NA) +
    geom_text_repel(data=subset(plot_data2, Patient_ID %in% c("SK-MEL-24", "HT-29","SK-MEL-30","UACC-62")),
                    aes(label=Label_Select, color=Label, segment.color=Label),
                    show.legend = FALSE,
                    direction = "x",
                    ylim = c(6, NA),
                    size=3) +
      geom_text_repel(data=subset(plot_data2,
                                  (Patient_ID %in% c("COLO-678","SK-MEL-2") &
                                    SimCase %in% c("RNA", "Mutations & RNA")) |
                                    Patient_ID %in% c("HT55","HT-144") &
                                    SimCase %in% c("RNA")),
                    aes(label=Label_Select, color=Label, segment.color=Label),
                    show.legend = FALSE,
                    direction = "y",
                    xlim = c(1.05, NA),
                    size=3) +
      geom_text_repel(data=subset(plot_data2,
                                  (Patient_ID %in% c("HT55","HT-144") &
                                     SimCase!="RNA") |
                                    (Patient_ID %in% c("COLO-678","SK-MEL-2") &
                                     SimCase=="Mutations")),
                    aes(label=Label_Select, color=Label, segment.color=Label),
                    show.legend = FALSE,
                    direction = "y",
                    xlim = c(NA, 0.45), ylim = c(NA, 2),
                    size=3) +
    ylim(c(-3,7)) + xlim(0, 1.3) +
    stat_cor(aes(label = ..r.label..), method="pearson", label.y.npc="bottom") +
    facet_grid(SimCase~Variant) +
    theme_pubclean() +
    scale_fill_manual(values = colors_types_4) +
    scale_color_manual(values = colors_types_4) +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 14),
          #strip.background.y = element_blank(),
          #strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7),
          legend.key = element_rect(fill = NA, colour = NA, size = 0.25),
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
    labs(y="CRISPR screening metric\n(scaled Bayesian factor)",
         #x=expression(paste("Proliferation"[BRAFinhibition]," / Proliferation"[control])),
         #x=TeX("$\\frac{Proliferation_{BRAFinhibited}}{Proliferation_{control}}$"),
         fill="Cancer type\n(and BRAF status)") +
      guides(color=FALSE, text=FALSE)
  
  
  p_2 <- egg::ggarrange(p_drugs_scatter2, p_crispr_scatter2,
                 nrow = 1, widths = c(4,4))

  #p_2

```

```{r scatter_pub2, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE, fig.height=6, fig.width=9}

plot_data <- filter(data_all_clin,
                    Target=="BRAF",
                    Metric %in% c("CC2", "AUC"),
                    Drug_Name!="Dabrafenib" | Screening=="CRISPR",
                    Output=="Prolif. norm.",
                    SimCase %in% sim_cases,
                    !is.na(Value)) %>%
  left_join(select(CL_mut_profile, Patient_ID, BRAF),
              by = "Patient_ID") %>%
    mutate(BRAF=case_when(
      BRAF==1 ~ "Gain function",
      BRAF==0 ~ "Loss function",
      is.na(BRAF) ~ "No mutation"
    ),
    SimCase=case_when(
      SimCase=="mut" ~ "Mutations",
      SimCase=="RNA" ~ "RNA",
      SimCase=="mut_RNA" ~ "Mutations & RNA"
    ) %>% factor(levels=c("Mutations", "RNA", "Mutations & RNA"))
    )
  
  p_drugs_scatter2 <- filter(plot_data, Screening=="Drug") %>%
    mutate(Label_Void=" ") %>%
    ggplot(aes(x=Score, y=Value)) +
    geom_label_repel(aes(label=Label_Void, fill=TCGA_label,
                         color=BRAF), label.size=0.75,
                     show.legend = F,
                     box.padding = 0, label.padding = 0.02,
                     segment.color = NA) +
    stat_cor(aes(label = ..r.label..),
             method="pearson", label.y.npc="top") +
    facet_grid(SimCase~Drug_Name) +
    theme_pubclean() +
    scale_fill_manual(values = colors_types) +
    scale_color_manual(values=c("Loss function"=NA, "Gain function"="#701B06FF")) +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 14),
          strip.background.y = element_blank(),
          strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7),
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
    labs(#x=expression(paste("Proliferation"[BRAFinhibition]," / Proliferation"[control]))#,
         #x=TeX("$\\frac{Proliferation_{BRAFinhibited}}{Proliferation_{control}}$"),
         y="Drug screening metric\n(AUC)"
         )
  
  plot_data2 <- filter(plot_data, Screening=="CRISPR") %>%
    mutate(Variant="Broad data", Label_Void=" ",
           Label_Select = if_else(Patient_ID %in% c("SK-MEL-24", "COLO-678", "HT-29", "HT-144", "HT55", "SK-MEL-30", "SK-MEL-2", "UACC-62"),
                                  Patient_ID, " "))
    p_crispr_scatter2 <- ggplot(plot_data2, aes(x=Score, y=Value)) +
    geom_label_repel(aes(label=Label_Void, fill=Label),
                     show.legend = TRUE,
                     box.padding = 0, label.padding = 0.02,
                     segment.color = NA) +
    geom_text_repel(data=subset(plot_data2, Patient_ID %in% c("SK-MEL-24", "HT-29","SK-MEL-30","UACC-62")),
                    aes(label=Label_Select, color=Label, segment.color=Label),
                    show.legend = FALSE,
                    direction = "x",
                    ylim = c(6, NA),
                    size=3) +
      geom_text_repel(data=subset(plot_data2,
                                  (Patient_ID %in% c("COLO-678","SK-MEL-2") &
                                    SimCase %in% c("RNA", "Mutations & RNA")) |
                                    Patient_ID %in% c("HT55","HT-144") &
                                    SimCase %in% c("RNA")),
                    aes(label=Label_Select, color=Label, segment.color=Label),
                    show.legend = FALSE,
                    direction = "y",
                    xlim = c(1.05, NA),
                    size=3) +
      geom_text_repel(data=subset(plot_data2,
                                  (Patient_ID %in% c("HT55","HT-144") &
                                     SimCase!="RNA") |
                                    (Patient_ID %in% c("COLO-678","SK-MEL-2") &
                                     SimCase=="Mutations")),
                    aes(label=Label_Select, color=Label, segment.color=Label),
                    show.legend = FALSE,
                    direction = "y",
                    xlim = c(NA, 0.45), ylim = c(NA, 2),
                    size=3) +
    ylim(c(-3,7)) + xlim(0, 1.3) +
    stat_cor(aes(label = ..r.label..), method="pearson", label.y.npc="bottom") +
    facet_grid(SimCase~Variant) +
    theme_pubclean() +
    scale_fill_manual(values = colors_types_4) +
    scale_color_manual(values = colors_types_4) +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 14),
          #strip.background.y = element_blank(),
          #strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7),
          legend.key = element_rect(fill = NA, colour = NA, size = 0.25),
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
    labs(y="CRISPR screening metric\n(scaled Bayesian factor)",
         #x=expression(paste("Proliferation"[BRAFinhibition]," / Proliferation"[control])),
         #x=TeX("$\\frac{Proliferation_{BRAFinhibited}}{Proliferation_{control}}$"),
         fill="Cancer type\n(and BRAF status)") +
      guides(color=FALSE, text=FALSE)
  
  
  p_2 <- egg::ggarrange(p_drugs_scatter2, p_crispr_scatter2,
                 nrow = 1, widths = c(4,4))

  #p_2

```

And here is the version with table

```{r scatter_pub_table, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.height=6, fig.width=12}

plot_data_tab <- filter(data_all_clin,
                    Target=="BRAF",
                    Metric == "CC2",
                    Output=="Prolif. norm.",
                    SimCase == "mut_RNA",
                    !is.na(Value)) %>%
    left_join(select(CL_mut_profile, Patient_ID, BRAF, p53, RAS, PI3K, MEK1_2),
              by = "Patient_ID")
  
  p_crispr_label <- mutate(plot_data_tab, Variant="Broad data") %>%
    ggplot(aes(x=Score, y=Value)) +
    geom_label_repel(aes(fill=TCGA_label, label=Patient_ID),
                     box.padding = 0, label.padding = 0.1,
                     size=4, color="white",segment.colour = NULL,
                     show.legend = FALSE) +
    theme_pubclean() +
    scale_fill_manual(values = colors_types) +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5),
          #strip.background.y = element_blank(),
          #strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7),
          legend.key = element_rect(fill = NA, colour = NA, size = 0.25),
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
    labs(y="CRISPR screening metric\n(scaled Bayesian factor)",
         x=TeX("$\\frac{Proliferation_{BRAFinhibited}}{Proliferation_{control}}$"),
         color="Cancer type",
         alpha="BRAF status")
  
  genes_mut <- c("BRAF", "RAS", "p53", "PI3K", "MEK1_2")
  genes_RNA <- c("CRAF", "ERK", "PDK1", "p70")
  CL_of_interest <- c("SK-MEL-24", "COLO-678", "HT-29", "HT-144", "HT55", "SK-MEL-30", "SK-MEL-2", "UACC-62")
  
  plot_data_basis <- filter(data_all_clin,
                    Target=="BRAF",
                    Metric == "CC2",
                    Output=="Prolif. norm.",
                    SimCase == "mut_RNA",
                    !is.na(Value)) %>%
    select(Patient_ID, TCGA_label)
  
  plot_data_mut <- left_join(plot_data_basis,
                             select(CL_mut_profile, Patient_ID,
                                    one_of(genes_mut)),
              by = "Patient_ID") %>%
      gather(any_of(genes_mut),key ="Model nodes", value="Value") %>%
    mutate(Type="Mutations profile\n(discrete)",
           `Model nodes`=if_else(`Model nodes`=="MEK1_2", "MEK",`Model nodes`))
  
  plot_data_RNA <- left_join(plot_data_basis,
                             select(CL_RNA_profile, Patient_ID,
                                    one_of(genes_RNA)),
              by = "Patient_ID") %>%
      gather(any_of(genes_RNA),key ="Model nodes", value="Value") %>%
    mutate(Type="RNA profile\n(continuous)",
           `Model nodes`=if_else(`Model nodes`=="PDK1", "PDPK1",`Model nodes`))
  
  plot_data_tab2 <- bind_rows(plot_data_mut, plot_data_RNA) %>%
    mutate(Value_char=as.character(Value)) %>%
    bind_rows(mutate(plot_data_basis, Value=0.5, `Model nodes`="",
                     Value_char=Patient_ID, Type="CL")) %>%
    mutate(Patient_ID=factor(Patient_ID, levels=rev(sort(my_cell_lines))),
           Alpha=if_else(is.na(Value), 0, 1),
           Size_char=if_else(Value_char %in% CL_of_interest, 4, 5)) %>%
    filter(Patient_ID %in% CL_of_interest)
   
   p_table_cl <- ggplot(plot_data_tab2,
                        aes(x=`Model nodes`, y=Patient_ID)) +
     geom_tile(color="white", aes(fill=Value, alpha=Alpha)) +
     geom_text(aes(label=Value_char, size=Size_char)) +
     facet_nested(.~Type, scales = "free", space = "free")+
     theme_pubclean() +
      theme(legend.position = "right",
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14),
            axis.ticks.y = element_blank(),
            axis.text.y = element_blank(),
            axis.title.y = element_blank(),
            #axis.text.x = element_text(size=8),
            #strip.background = element_blank(),
            strip.text = element_text(size=14)#,
            #panel.grid.major.y = element_blank()
        ) +
     scale_fill_paletteer_c("ggthemes::Classic Red-White-Green",
                            limits=c(0, 1)) +
     scale_radius(range = c(2.5,5)) +
     guides(alpha=FALSE, size=FALSE) +
     labs(fill="Node value")
  
  p_3 <- egg::ggarrange(p_crispr_label, p_table_cl,
                 nrow = 1, widths = c(4,4))

  #p_3
  
  #ggsave(file="p_1.svg", plot=p_1, width=12, height=3)
  #ggsave(file="p_2.svg", plot=p_2, width=12, height=7.8)
  #ggsave(file="p_3.svg", plot=p_3, width=12, height=5.4)

```

Additional plot for p53 and PI3K:

```{r scatter_pub_add, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.height=6, fig.width=10}
colors_types_add <- c("p53: Loss function"="#B18147FF",
                  "p53: No mutation"="#B2AAA2FF",
                  "PI3K: Gain function"="#803B31FF",
                  "PI3K: No mutation"="#B2AAA2FF")

plot_data <- filter(data_all_clin,
                    Target %in% c("PI3K", "p53"),
                    Metric %in% c("CC2"),
                    Output=="Prolif. norm.",
                    SimCase %in% c("mut", "mut_RNA"),
                    !is.na(Value)) %>%
  left_join(select(CL_mut_profile, Patient_ID, BRAF, p53, PI3K),
              by = "Patient_ID") %>%
  mutate(p53=case_when(
      p53==1 ~ "p53: Gain function",
      p53==0 ~ "p53: Loss function",
      is.na(p53) ~ "p53: No mutation"
    ),
    PI3K=case_when(
      PI3K==1 ~ "PI3K: Gain function",
      PI3K==0 ~ "PI3K: Loss function",
      is.na(PI3K) ~ "PI3K: No mutation"
    ),
    SimCase=case_when(
      SimCase=="mut" ~ "Mutations",
      SimCase=="RNArb" ~ "RNA",
      SimCase=="mut_RNArb" ~ "Mutations & RNA"
    ) %>% factor(levels=c("Mutations", "RNA", "Mutations & RNA"))
    )

p_table_red <- select(plot_data, SimCase, P_Mutants, P_Rates) %>%
   pivot_longer(starts_with("P_"), names_to="P_Method", values_to="P_Type") %>%
   distinct %>%
   mutate(Alpha_Status=if_else(P_Type==" ", 0, 1),
         P_Method=if_else(P_Method=="P_Mutants", "Discrete", "Continuous") %>% factor(levels=c("Discrete", "Continuous")),
         Label="Personalization method") %>%
   ggplot(aes(x=P_Method, y=1)) +
      geom_tile(aes(alpha=as.factor(Alpha_Status)),
                color="black", size=0.5,fill="grey85",
                show.legend = FALSE) +
      geom_text(aes(label=P_Type), color="black", angle=90) +
      theme_pubclean() +
      theme(panel.grid.minor.y = element_blank(),
            strip.text = element_text(size=12),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank(),
            axis.text.y = element_blank(),
            strip.background.y = element_blank(),
            strip.text.y = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.x = element_blank()) +
      scale_fill_manual(values=colors_perso) +
      scale_alpha_manual(values = c(0.4,1)) +
  scale_y_continuous(breaks = seq(1, 1, 1)) +
  labs(x="Personalization methods",
       y="Combinations of personalization methods") +
  facet_nested(SimCase~Label+P_Method, scales = "free", space = "free")
  
  p_crispr_p53 <- filter(plot_data, Target=="p53") %>%
    mutate(Variant="Broad CRISPR data: TP53 inhibition", Label_Void=" ") %>%
    ggplot(aes(x=Score, y=Value)) +
    geom_label_repel(aes(label=Label_Void, fill=p53),
                     show.legend = TRUE,
                     box.padding = 0, label.padding = 0.03,
                     segment.color = NA) +
    stat_cor(method="pearson",
             label.y.npc=0.2, label.sep = "\n") +
    facet_grid(SimCase~Variant) +
    theme_pubclean() +
    scale_fill_manual(values = colors_types_add) +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          strip.background.y = element_blank(),
          strip.text = element_text(size = 12),
          strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7),
          legend.key = element_rect(fill = NA, colour = NA, size = 0.25),
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
    labs(y="CRISPR screening metric\n(scaled Bayesian factor)",
         #x=expression(paste("Proliferation"[BRAFinhibition]," / Proliferation"[control])),
         x=TeX("$\\frac{Proliferation_{p53inhibited}}{Proliferation_{control}}$"),
         fill="Mutation status: ") +
    guides(fill=guide_legend(nrow=1,byrow=TRUE))
  
  p_crispr_PI3K <- filter(plot_data, Target=="PI3K") %>%
    mutate(Variant="Broad CRISPR data: PIK3CA inhibition", Label_Void=" ") %>%
    ggplot(aes(x=Score, y=Value)) +
    geom_label_repel(aes(label=Label_Void, fill=PI3K),
                     show.legend = TRUE,
                     box.padding = 0, label.padding = 0.03,
                     segment.color = NA) +
    stat_cor(method="pearson",
             label.y.npc=0.2, label.sep = "\n") +
    facet_grid(SimCase~Variant) +
    theme_pubclean() +
    scale_fill_manual(values = colors_types_add) +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size=12),
          strip.background.y = element_blank(),
          strip.text.y = element_blank(),
          panel.border = element_rect(color = "gray30",
                                      fill = NA, size = 0.7),
          legend.key = element_rect(fill = NA, colour = NA, size = 0.25),
          legend.title = element_blank(),
          axis.title.y = element_blank(),
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14)) +
    labs(#x=expression(paste("Proliferation"[BRAFinhibition]," / Proliferation"[control])),
         x=TeX("$\\frac{Proliferation_{PI3Kinhibited}}{Proliferation_{control}}$"),
         fill="Cancer type\n(and BRAF status)") +
    guides(fill=guide_legend(nrow=1,byrow=TRUE))
  
  
  p_add <- egg::ggarrange(p_table_red, p_crispr_p53, p_crispr_PI3K,
                 nrow = 1, widths = c(2,4,4))

  p_add

  #ggsave(file="p_add.svg", plot=p_add, width=12, height=6)
```

## Interactive plots

We can have a deeper look at scatter plot with interactive settings

Here is the non-interactive reference plots

```{r scatter_reference, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.height=6, fig.width=9}
p_2
```

Let's generate each column as an interactive plot, first with drugs and then with CRISPR:

```{r scatter_interactive, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.height=6, fig.width=9}

p_table_interactive <- select(corr_red, SimCase, P_Mutants, P_Rates) %>%
  pivot_longer(starts_with("P_"), names_to="P_Method", values_to="P_Type") %>%
  mutate(Alpha_Status=if_else(P_Type==" ", 0, 1),
         P_Method=if_else(P_Method=="P_Mutants", "Discrete", "Continuous"),
         SimCase=factor(SimCase, levels=sim_cases)) %>%
  distinct %>%
  ggplot(aes(x=P_Method, y=1)) +
  geom_tile(aes(fill=as.factor(Alpha_Status)),
            color="black", size=0.5,
            show.legend = FALSE) +
  geom_text(aes(label=P_Type), color="black") +
  theme_pubclean() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position='none') +
  scale_fill_manual(values = c("0"="grey90", "1"="grey50")) +
  labs(x="Personalization methods",
       y="Combinations of personalization methods") +
  facet_grid(SimCase~.)
ply_table_interactive <- ggplotly(p_table_interactive, tooltip = "x")

#Panel Drugs
p_drugs <- filter(data_all_clin_reduced,
                  Output=="Prolif. norm.", Screening=="Drug", Drug_Name=="PLX-4720") %>%
  ggplot(aes(x=Score, y=Value)) +
  geom_point(aes(color=TCGA_label, text=Label),
             show.legend = FALSE) +
  facet_grid(SimCase~.) +
  theme_pubclean() +
  scale_color_manual(values = colors_types)  +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        strip.background.y = element_blank(),
        strip.text.y = element_blank(),
        panel.border = element_rect(color = "gray30",
                                    fill = NA, size = 0.7)) +
  labs(title="Drug: (Prolif. norm.) VS (PLX-4720 AUC)")

ply_drugs <- ggplotly(p_drugs, tooltip = "text")
subplot(p_table_interactive, ply_drugs, widths =c(0.3, 0.7))

#Panel B
p_crispr <- filter(data_all_clin_reduced,
                   Output=="Prolif. norm.",
                   Screening=="CRISPR",
                   Metric=="CC2", Target=="BRAF") %>%
  ggplot(aes(x=Score, y=Value)) +
  geom_point(aes(color=TCGA_label, text=Label),
             show.legend = FALSE) +
  facet_grid(SimCase~.) +
  theme_pubclean() +
  scale_color_manual(values = colors_types)  +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        strip.background.y = element_blank(),
        strip.text.y = element_blank(),
        panel.border = element_rect(color = "gray30",
                                    fill = NA, size = 0.7)) +
  labs(title="CRISPR: (Prolif. norm.) VS (CRISPR Broad)")
ply_crispr <- ggplotly(p_crispr, tooltip = "text")

subplot(p_table_interactive, ply_crispr, widths =c(0.3, 0.7))
```

And a last interactive plot to visualize the benefit of RNA addition for CRISPR prediction:

```{r, multi_omics, echo=FALSE, warning=FALSE, fig.width = 10, fig.height=4.5, eval=TRUE}

plot_data <- filter(data_all_clin,
                    Target=="BRAF",
                    Metric %in% c("CC2"),
                    P_Mutants=="Mutations",
                    Output=="Prolif. norm.",
                    SimCase %in% c('mut', 'mut_RNA'),
                    !is.na(Value)) %>%
  mutate(Label=paste0(Patient_ID, " (", TCGA_label,")\n",Label_Mut, "\n", Label_RNA)) %>%
  mutate(SimCase=if_else(SimCase=="mut", "Mutations", "Mutations & RNA"))

p_crispr <- ggplot(plot_data, aes(x=Score, y=Value)) +
  geom_point(aes(color=TCGA_label, text=Label, frame=SimCase),
             show.legend = FALSE) +
  #facet_grid(SimCase~.) +
  theme_pubclean() +
  scale_color_manual(values = colors_types)  +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        strip.background.y = element_blank(),
        strip.text.y = element_blank(),
        panel.border = element_rect(color = "gray30",
                                    fill = NA, size = 0.7)) +
  labs(title="CRISPR: (Prolif. norm.) VS (CRISPR Broad)",
       x="Score from model",
       y="Exp. sensitivity")
ply_crispr <- ggplotly(p_crispr, tooltip = "text") %>%
  animation_slider(currentvalue = list(prefix = "Personalization with ", font = list(color="black")),
                   lenmode="fraction",
                   len=0.5,
                   x=0.2,
                   y=-0.1)

ply_crispr

```
