---
title: "Comparison_ML"
author: "Jonas BÉAL"
date: "30/03/2020"
output:
  html_document:
    number_sections: yes
---

This notebook focuses on machine learning (ML) approaches. The idea is not to compare performances with logical models but to answer the following questions:

* Are the variables selected by ML alorithms the same as the ones used by personalized logical models?
* What kind of interpretation could we derive from simple ML analysis?

Since the corresponding publication is not focused on ML, we will not benchmark all possible ML approaches or perfectly fine-tune all the parameters. The objective is to compare with a simple "out-of-the-box" ML approach. We chose to focus on random forests for their versatility and existing post-hoc interpretability methods (variable importance etc.)

```{r setup, include=FALSE}
#Infos from simulations
model <- c("BRAF_Model")
models_outputs <- c("Proliferation_b1","Proliferation_b2")

sim_cases <- c("mut","RNA", "mut_RNA")

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#Packages
if (!require("pacman")) install.packages("pacman")
list.of.packages <- c("readxl","knitr","corrplot","ggfortify","reshape2","gridExtra","magrittr", "tictoc", "ggrepel", "broom", "nlme", "lme4", "lmerTest", "eulerr", "plotly", "paletteer", "ggpubr", "cowplot", "ggcorrplot", "tidyverse", "latex2exp", "randomForestSRC", "tictoc")
pacman::p_load(list.of.packages, character.only = TRUE)
select <- dplyr::select
tbl_to_df <- function(x){
  x <- x %>% as.data.frame %>%  remove_rownames %>% column_to_rownames(var="model_name")
}

#Paths
original_path <- normalizePath(getwd())
base_folder <- original_path %>% dirname %>% dirname
knitr::opts_knit$set(root.dir = base_folder)
```

# Import data

Let's import the following data :

* Mutations and RNA data directly from data files (original data)
* Mutations and RNA data from the logical model profiles (processed data)
* Sensitivity metrics (drug and CRISPR)

```{r DataImport, echo=FALSE, message=FALSE, warning=FALSE}
tic()

#Genes involved in PC model
genenames <- read.table(paste0("Models/",model,"/",model,"_namesToHugo_curated.txt"),
                         header=T,sep="\t")
geneindex <- strsplit(as.character(genenames[,2]), split = ",") %>% sapply(function(l){gsub(" ","",l)})
geneindex <- data.frame(V1 = rep(genenames[,1], sapply(geneindex, length)), V2 = unlist(geneindex))
model_genes_HUGO <- unique(geneindex$V2) %>% sub("^\\s+", "", .)


#First import CL clinical data for further use to correlate with logical modelling results
CL_clin <- read_csv("Data/CL/Infos/model_list_20200115.csv") %>%
  select(model_id, model_name, synonyms, model_type, COSMIC_ID, BROAD_ID,
         contains("_data"), contains("tissue"), contains("cancer")) %>%
  filter(mutation_data | cnv_data | expression_data) %>%
  mutate(TCGA_label=if_else(cancer_type=="Colorectal Carcinoma", "CRC",
                            if_else(cancer_type=="Melanoma", "CM", cancer_type)))

#Focus on two cancer types only
good_labels <- c("CRC", "CM")
CL_clin %<>% filter(TCGA_label %in% good_labels)
my_cell_lines <- unlist(CL_clin$model_name)
my_cell_lines_s <- unlist(CL_clin$model_id)

#Fonction to transpose some tibbles
tibble_transpose <- function(df_input){
  df_output <- df_input %>% gather(var, value, -gene_symbol) %>%
    spread(gene_symbol, value) %>%
    rename(model_name=var) %>% 
    type_convert
}

#Import CL mutations profile
CL_mut_processed <- read_csv(paste0("Results/Profiles/",model,"_CL_mut.csv")) %>%
  rename(Patient_ID=X1) %>%
  select(-matches("_b2")) %>%
  rename_all(function(x) gsub("_b1", "", x)) %>%
  mutate_at(vars(-Patient_ID), function(x) if_else(!is.na(x), 1, 0)) %>%
  rename_at(vars(-Patient_ID), function(x) paste0("Mut_",x)) %>%
  mutate_at(vars(-Patient_ID), as.factor)

#Import CL RNA profile
CL_RNA_processed <- read_csv(paste0("Results/Profiles/",model,"_CL_RNA_norm.csv")) %>%
  rename(Patient_ID=X1) %>%
  select(-matches("_b2")) %>%
  rename_all(function(x) gsub("_b1", "", x)) %>%
  rename_at(vars(-Patient_ID), function(x) paste0("RNA_",x))

CL_processed <- full_join(CL_mut_processed, CL_RNA_processed, by="Patient_ID")

#Import raw data
#Select cancer genes to reduce size
cancer_genes_cosmic <- read_delim("Data/Common/cancer_genes_COSMIC.txt", delim = ',') %>% unlist 
cancer_genes_HPA <- read_delim("Data/Common/cancer_genes_HPA.txt", delim = ',') %>% unlist 
cancer_genes <- c(cancer_genes_cosmic, cancer_genes_HPA) %>% unique

#Import CL raw mutations data
CL_mut_original <- read_csv("Data/CL/mutations_20191101.csv") %>% 
  filter(model_name %in% my_cell_lines, gene_symbol %in% cancer_genes) %>%
  select(model_name, gene_symbol) %>%
  mutate(Status=1) %>% distinct %>%
  pivot_wider(names_from=gene_symbol, values_from=Status, values_fill = list(Status = 0)) %>%
  rename_at(vars(-model_name), function(x) paste0("Mut_",x)) %>%
  mutate_at(vars(-model_name), as.factor)

#Import CL raw RNA data
CL_RNA_original <- read_csv("Data/CL/rnaseq_fpkm_20191101.csv", skip=1) %>%
  rename(gene_symbol=X2) %>% slice(-c(1,2)) %>%
  filter(gene_symbol %in% c(cancer_genes, model_genes_HUGO)) %>%
  select(gene_symbol, one_of(my_cell_lines)) %>%
  tibble_transpose %>%
  mutate_if(is_numeric, function(x) log2(x+1)) %>%
  rename_at(vars(-model_name), function(x) paste0("RNA_",x))

CL_original <- full_join(CL_mut_original, CL_RNA_original, by="model_name") %>%
  rename(Patient_ID=model_name)

CL_original_red <- select(CL_original, Patient_ID,
                          any_of(lapply(c("Mut_", "RNA_"),
                                        function(x) paste0(x, model_genes_HUGO)) %>% unlist))

#Import GDSC drug screening
Drugs_Profile_Conversion <- read_csv(paste0("Results/Profiles/", model, "_drug_conversion.csv")) %>%
  rename(Drug_ID=DRUG_ID, Drug_Name=DRUG_NAME) %>%
  mutate(Drug_ID=as.character(Drug_ID))

#Prepare conversion
DC <- Drugs_Profile_Conversion %>%
  mutate(Variant=1)
for (r in 2:nrow(DC)){
  prof <- DC$Drug_Target[r]
  DC[r,"Variant"] <- 1+sum(DC$Drug_Target[1:(r-1)]==prof)
}
DC %<>% mutate(Variant=paste0(Drug_Target, "_",Variant)) %>%
  add_row(Drug_ID=NA,Drug_Target="Zero",
          Drug_Name="Zero", Variant="Zero") %>%
  rename(Target=Drug_Target)


#BRAF_drugs <- c("PLX-4720", "Dabrafenib")
CL_Drugs <- read_xlsx("Data/CL/GDSC2_fitted_dose_response_15Oct19.xlsx") %>%
  filter(SANGER_MODEL_ID %in% my_cell_lines_s) %>%
  mutate(CELL_LINE_NAME=if_else(CELL_LINE_NAME=="LS-1034", "LS1034", CELL_LINE_NAME)) %>%
  rename(Patient_ID=CELL_LINE_NAME, Drug_Name=DRUG_NAME) %>%
  left_join(rename(CL_clin, Patient_ID=model_name) %>% select(Patient_ID, TCGA_label),
            by="Patient_ID") %>%
  filter(Drug_Name %in% DC$Drug_Name) %>%
  select(Patient_ID, Drug_Name, Z_SCORE, AUC, TCGA_label) %>%
  left_join(select(DC, -Drug_ID), by="Drug_Name") %>%
  pivot_longer(one_of("Z_SCORE", "AUC"), names_to="Metric", values_to="Value")
  
#Import CRISPR/Cas9 screening
CL_CC1 <- read_tsv("Data/CL/03_scaledBayesianFactors.tsv") %>%
  filter(Gene %in% model_genes_HUGO) %>%
  rename(LS1034=`LS-1034`) %>%
  select(Gene, one_of(my_cell_lines)) %>%
  pivot_longer(-Gene, names_to="Patient_ID", values_to="Value") %>%
  left_join(rename(CL_clin, Patient_ID=model_name) %>% select(Patient_ID, TCGA_label),
            by="Patient_ID")

CL_CC2 <- read_tsv("Data/CL/scaledBayesianFactors.tsv") %>%
  filter(Gene %in% model_genes_HUGO) %>%
  pivot_longer(-Gene, names_to="Patient_ID", values_to="Value") %>%
  rename(BROAD_ID=Patient_ID) %>%
  left_join(select(CL_clin, BROAD_ID, model_name, TCGA_label),
            by="BROAD_ID")  %>%
  mutate(model_name=if_else(model_name=="LS-1034", "LS1034", model_name)) %>%
  rename(Patient_ID=model_name) %>%
  filter(Patient_ID %in% my_cell_lines) %>%
  select(-BROAD_ID)

CL_CC_All <- bind_rows(mutate(CL_CC1, Metric="CC1"),
                       mutate(CL_CC2, Metric="CC2")) %>%
  rename(Variant=Gene) %>%
  left_join(rename(geneindex, Target=V1, Variant=V2),
            by="Variant")

#Merge all metric
CL_Metrics <- bind_rows(mutate(CL_Drugs, Screening="Drug"),
                        mutate(CL_CC_All, Screening="CRISPR"))
  
CL_Metrics_Sample_Size <- CL_Metrics %>%
  filter(!is.na(Value)) %>%
  group_by(TCGA_label, Screening, Metric, Variant) %>%
  summarize(Size=n()) %>%
  ungroup()

#Nodes data
import_ALL <- function(sim_case){
  simulations <- read_table2(paste0("Results/Simulations/",model,"_CL_",sim_case,".txt")) %>%
    select(-Time, -n_profile) %>%
    mutate(Proliferation=(Proliferation_b1 + Proliferation_b2)/2)
  return(simulations)
}

simulations <- sapply(sim_cases, import_ALL, simplify = F, USE.NAMES = T)


print("All imports OK")
```

Please note that, to reduce the size of mutations and RNA matrices, we focus only on cancer-related genes using a merged list from COSMIC and HPA (Human Proteome Atlas) composed of `r length(cancer_genes)`. So all in all we build 3 different matrices of input data:

* Matrix **CL_processed** includes mutations and RNA from profiles used for logical models personalization
* Matrix **CL_original** includes mutations and RNA only for cancer-related genes (`r length(cancer_genes)`) and model-related genes
* Matrix **CL_original_red** includes mutations and RNA only model-related genes (`r length(model_genes_HUGO)`)

# Random forest implementation

To increase the number of samples and favour the learning of RF algorithms we decided to merge similar learning task:

* For **Drugs** screening, outcome of BRAF inhibition through PLX and Dabrafenib are conidered altogether (using z-scores $log(IC50)$ to reinforce comparability between drugs)
* For **CRISPR** screening, datasets from Broad and Sanger screenings are merged

In both cases this strategy results in duplicate patients with potentially different outcomes

```{r rf_compute, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}

rf_performances <- data.frame(Screening=NA_character_,
                              Data=NA_character_,
                              Processing=NA_character_,
                              Exp_Var=NA_real_,
                              N_Var=NA_real_,
                              N_Samples=NA_real_) %>%
  filter(!is.na(Screening))

# CL_Metrics_forMerge <- filter(CL_Metrics,
#                               Target=="BRAF",
#                               (Screening=="Drug" & Metric=="Z_SCORE" & Drug_Name=="PLX-4720") |
#                                 (Screening=="CRISPR" & Metric=="CC2")) %>%
#   select(-Drug_Name, -TCGA_label, -Target, -Variant, -Metric)

CL_Metrics_forMerge <- filter(CL_Metrics,
                              Target=="BRAF",
                              (Screening=="Drug" & Metric=="Z_SCORE") |
                                (Screening=="CRISPR")) %>%
  select(-Drug_Name, -TCGA_label, -Target, -Variant, -Metric)

for (s in c("Drug", "CRISPR")){
  for (d in c("Mut", "RNA", "Mut_RNA")){
    for (p in c("original", "original_red", "processed")){
      data_rf <- full_join(CL_Metrics_forMerge,
                           get(paste0("CL_", p)),
                           by="Patient_ID") %>%
        filter(!is.na(Value),
               Screening==!!s) %>%
        select(-Patient_ID, -Screening) %>%
        as.data.frame
      
      if (d != "Mut_RNA"){
        data_rf %<>% select(Value, starts_with(d))
      }
      
      fit <- rfsrc(Value~., data = data_rf, importance = T)
      
      exp_var <- 100*(1 - ( mean( (fit$yvar-fit$predicted.oob)^2 ) / var(fit$yvar) ))
      
      imp <- fit$importance %>% sort(decreasing = T) %>% head(n=10) %>%
        enframe(name="Variable", value = "Importance") %>%
        mutate(Screening=s, Data=d, Processing=p, Exp_Var=exp_var,
               N_Var=ncol(data_rf)-1,
               N_Samples=sum(rowSums(is.na(data_rf)) < ncol(data_rf)-1))
        
      # rf_performances %<>% add_row(Screening=s, Data=d, Processing=p,
      #                              Exp_Var=exp_var)
      
      rf_performances %<>% bind_rows(imp)
      
    }
  }
}

#i <- find.interaction(t)
print("RF fitting done")
```

$100 \times (1- \dfrac{Mean[(Y_{observed}-Y_{predicted})^2]}{Var(Y_{observed})})$

## Percentage of variance explained

We use radom forests to predict sensitivity to BRAF inhibition (with drugs or CRISPR). Let's compare the percentage of variance explained by the fitted random forests in different cases. Each learning task is a combination of:

* the processing of input dataset (original data, original data for model genes only, processed data for model profiles)
* the kind of omics data used (mutatios only, RNA only, mutations & RNA)
* the nature of outcome predicted (from drug screening or CRISPR screening).

```{r rf_plot, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}

data_colors <- c("Mut"="#306489FF", "RNA"="#EBCB8BFF", "Mut_RNA"="#A3BE8CFF")


p_expvar <- select(rf_performances, -Variable, -Importance) %>% distinct %>%
  mutate(Processing=case_when(
         Processing=="original" ~ "Original data\n(cancer and model genes)",
         Processing=="original_red" ~ "Original data\n(model genes only)",
         Processing=="processed" ~ "Processed data\n(model profiles)"
       )) %>%
  mutate(Screening=factor(Screening, levels=c("Drug", "CRISPR")),
         Data=factor(Data, levels=c("Mut", "RNA", "Mut_RNA")),
         Label=paste0("n=",N_Samples , "\np=", N_Var)) %>%
  ggplot(aes(x=Data, y=Exp_Var)) +
  geom_bar(stat = "identity", aes(fill=Data)) +
  geom_text(aes(label=Label), size=4, color="white", y=13, angle=0) +
  scale_fill_manual(values=data_colors) +
  facet_grid(Screening~Processing) +
  theme_pubclean() +
  guides(fill=FALSE) +
  labs(title = "Performances of random forests for BRAF sensitivity prediction",
       subtitle = "(n samples and p variables per learning task)",
       x = "Omics data used",
       y = "% of explained variance")

p_expvar
```

```{r rf_plot_pub, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
ggsave(file="p_expvar.svg", plot=p_expvar, width=12, height=6)
```

We can derive some insights from the above plot :

* Original and processed data result in very similar performances for Drugs sensitivity prediction. Nevertheless, for CRISPR sensitivity prediction, performances increased when we use only model genes and even more when we use direcly model profiles. One hypothesis: since we have less samples in CRISPR data, reducing the number of variables may have a positive influence. This effect is not present in Drugs but we have 2 times more samples. But all in all, we can say that our data processing before model personalization does not seem to significantly decrease the amount of information in the data.

* For Drug sensitivity prediction mutation and RNA data (separately) seem to result in similar performances. But merging them together does not provide significantly improved results. For CRISPR sensitivity prediction, RNA data seems less informative, especially with reduced datasets. Merging mutations and RNA does not improve drastically overall performance.


## Variable importance

But beyond performance we are mainly interested by interpretation so we will have a look at Variable Importance (VIMP) as measured by the random forests.

```{r rf_vimp, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}


p_varimp <- filter(rf_performances, Screening=="Drug", Importance>0) %>%
  mutate(Processing=case_when(
         Processing=="original" ~ "Original data\n(cancer and model genes)",
         Processing=="original_red" ~ "Original data\n(model genes only)",
         Processing=="processed" ~ "Processed data\n(model profiles)"
       )) %>%
  mutate(#Screening=factor(Screening, levels=c("Drug", "CRISPR")),
         Data=factor(Data, levels=c("Mut", "RNA", "Mut_RNA"))) %>%
  ggplot(aes(x=reorder(Variable, -Importance), y=Importance)) +
  geom_bar(stat = "identity", aes(fill=Data)) +
  scale_fill_manual(values=data_colors) +
  facet_grid(Data~Processing, scales = "free", space = "free") +
  theme_pubclean() +
  theme(axis.text.x = element_text(size=6, angle = 90)) +
  guides(fill=FALSE) +
  labs(title = "Random forests for BRAF drug sensitivity - Variable importance",
       subtitle = "(10 best variables with positive importance for each case)",
       x = "Variable",
       y = "Variable importance")
p_varimp

filter(rf_performances, Screening=="CRISPR", Importance>0) %>%
  mutate(Processing=case_when(
         Processing=="original" ~ "Original data\n(cancer and model genes)",
         Processing=="original_red" ~ "Original data\n(model genes only)",
         Processing=="processed" ~ "Processed data\n(model profiles)"
       )) %>%
  mutate(#Screening=factor(Screening, levels=c("Drug", "CRISPR")),
         Data=factor(Data, levels=c("Mut", "RNA", "Mut_RNA"))) %>%
  ggplot(aes(x=reorder(Variable, -Importance), y=Importance)) +
  geom_bar(stat = "identity", aes(fill=Data)) +
  scale_fill_manual(values=data_colors) +
  facet_grid(Data~Processing, scales = "free", space = "free") +
  theme_pubclean() +
  theme(axis.text.x = element_text(size=6, angle = 90)) +
  guides(fill=FALSE) +
  labs(title = "Random forests for BRAF CRISPR sensitivity - Variable importance",
       subtitle = "(10 best variables with positive importance for each case)",
       x = "Variable",
       y = "Variable importance")

```

```{r rf_plot_pub2, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
ggsave(file="p_varimp.svg", plot=p_varimp, width=12, height=6)
```

Mutational status of BRAF is the main driver of sensitivity prediction. RAS and p53 also appear as important mutations, taken into account in the model. The APC gene is the main important mutational variable not present in the model.

## Variable interactions

Now let's visualize the synergies between variables (by pairs). We focus here on Drugs screening with original data restricted to model nodes, using both mutations and RNA. On the following plot, off-diagonal coefficients show the synergistic effect of the combination of variable (). Diagonal coefficients remind the variable importance of variable alone.

```{r rf_interaction, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}

plot_interactions <- function(dataset){

data_int <- full_join(CL_Metrics_forMerge, get(paste0("CL_", dataset)),
                      by="Patient_ID") %>%
  filter(!is.na(Value),
         Screening=="Drug") %>%
  select(-Patient_ID, -Screening) %>%
  as.data.frame
      
fit <- rfsrc(Value~., data = data_int, importance = "permute")

fit_var_imp <- fit$importance %>% sort(decreasing = T) %>% head(n=10) %>% names

fit_int <- find.interaction(object =fit, xvar.names = fit_var_imp,
                            importance = "permute", method = "vimp",
                            verbose = F) 

plot_data <- fit_int %>% as_tibble(rownames = NA) %>%
  select(Difference) %>%
  rownames_to_column(var="Pair") %>%
  mutate(Row=strsplit(Pair, ":") %>% lapply(`[[`, 1) %>% unlist,
         Column=strsplit(Pair, ":") %>% lapply(`[[`, 2) %>% unlist) %>%
  select(-Pair) %>%
  bind_rows(mutate(., I=Column, Column=Row) %>% mutate(Row=I)) %>% select(-I) %>%
  rename(VIMP=Difference)
  #wider(names_from=Column, values_from=Difference) %>%
  #select(Row, one_of(.$Row))

indiv_vimp <- fit$importance %>% enframe(name = "Row", value = "VIMP") %>%
  filter(Row %in% fit_var_imp) %>%
  mutate(Column=Row) %>%
  arrange(desc(VIMP))

plot_data %<>% #bind_rows(indiv_vimp) %>%
  mutate(VIMP=if_else(VIMP<0, 0, VIMP),
         Row=factor(Row, levels=indiv_vimp$Row),
         Column=factor(Column, levels=indiv_vimp$Column))

ggplot(plot_data, aes(x=Column, y=Row, VIMP)) +
  geom_tile(color="black", aes(fill=VIMP)) +
  geom_text(data=indiv_vimp, aes(label=round(VIMP, digits=2))) +
  scale_fill_paletteer_c("pals::ocean.tempo", direction = +1, limits=c(0, NA)) +
  #scale_fill_paletteer_c("scico::lapaz", direction = -1, limits=c(0, NA)) +
  theme_pubclean() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        panel.grid.major.y = element_blank()) +
  #scale_x_discrete(position = "top") +
  labs(fill="Variable\nimportance",
       x="",
       y="",
       title = paste0("Importance of interactions between variables\n(", dataset ," dataset)"))

}

plot_interactions("processed")
plot_interactions("original_red")
plot_interactions("original")

```

We mainly observed important interaction/synergies between BRAF and RAS mutational status. But also between BRAF and some RNA levels, especially FOXD3 and SPRY. 

```{r eval=TRUE, echo=FALSE}
print("Done")
print(paste0("Computation time:"))
toc()
```
